Evaluating the Capabilities of Lattice Boltzmann Method for non-Newtonian and Free-surface Flows towards Applications in Wellbore Cementing

### 2016-07-26 thesis.tex
fixed a typo
fixed mathematical error
Chapter 1 first round of revisions ready for review

Analytical solutions rarely exist for even the simplest non-Newtonian flows because of the complexity that a nonlinear constitutive relationship entails.
It is generally more practical to approximate non-Newtonian flows using numerical methods. %MG: ask John about the following pargraph...
However, the nonlinear constitutive equation--typically of the form \begin{equation} \label{eq:non-newtonian}
\tau = \mu_{app}(\dot{\gamma}) \dot{\gamma},
\end{equation}\noindent where the apparent viscosity, $\mu_{app}$ is a function of the strain rate--results in certain challenges for numerical methods as well.
Determining the apparent viscosity and shear rate of a flow will often require an iterative solution, a general Picard-type algorithm is:

@@

A criteria for introducing artificial dissipation that has been used successfully is to define a threshold, $\theta$, such that dissipation is added when $\Delta S > \theta$~\cite{gorban2014enhancement,brownlee2006stabilization,brownlee2007stability,brownlee2008nonequilibrium}.
A potential drawback of defining a threshold \textit{a priori} is that in order to ensure the model still retains some physical integrity, only a small number of sites can have artificial dissipation added.
If the threshold is too low, too many sites may have dissipation added.
If the threshold is too high, a stable solution may not be achieved.
The threshold can be determined on a case-by-case basis through trial-and-error or by a preliminary analysis.
In the current work, the criteria that is used for determining whether dissipation should be added is:

\begin{equation} \label{eq:dissipation-criteria}
    \Delta S > \overline{\Delta S} + n_\sigma \cdot \sigma_{\Delta S},
\end{equation}

\noindent where $\overline{\Delta S}$ and $\sigma_{\Delta S}$ are the mean and standard deviation of $\Delta S$, respectively--both are calculated using values over the domain for the current timestep--and $n_\sigma$ is the number of standard deviations greater than $\overline{\Delta S}$ that $\Delta S$ must be before dissipation is added.
The number of standard deviations, $n_\sigma$, is chosen \textit{a priori}.
The criteria described in \eqref{eq:dissipation-criteria} has the advantage that one does not need to determine \textit{a priori} what constitutes ``far'' from quasiequilibrium, but instead can think in terms of what would be the maximum percentage of sites one would want artificial dissipation to be added to.

@@

In order to solve for the apparent viscosity, $\mu_{app}$, using the Picard-type algorithm outlined in \Fref{sec:lbm-for-nnf}, one must calculate the strain-rate, $\dot{\gamma}$.
The strain-rate is given by the second invariant of the strain-rate tensor, $\srtensor$, i.e.:
\begin{equation} \label{eq:second-invariant}
\dot{\gamma} = \sqrt{2 \sum_{\alpha,\beta=1}^{2} \srtensor}.
\end{equation}

@@

A numerical study was carried out in order to investigate the suitability of different LBM collision schemes for simulating non-Newtonian flows in terms of accuracy, numerical stability and computational efficiency.
For all simulations presented in this section, the apparent viscosity is approximated using its corresponding constitutive relationship and the Picard-type algorithm outlined in \Fref{sec:lbm-for-nnf} with the maximum number of iterations set at 15 and the convergence criteria:
\begin{equation} \label{eq:mu-app-conv}
\frac{\left|\mu_{app}^{k+1} - \mu_{app}^{k}\right|}{\mu_{app}^{k}} < 1.0 \times 10^{-6}.
\end{equation}
All BGK-BRT collision schemes use $\omega \in [0.05, 1.995]$ as the bounds.
All simulations with artificial dissipation use the median filter with $n_s = 2.7$.
All numerical values given in this section are in lattice units unless otherwise stated.

\newcommand{\pgrad}{\frac{\partial p}{\partial x}}

@@

Poiseuille flow is a useful benchmark because analytical solutions exist for both Bingham plastic and power-law fluids.
Poiseuille flow is driven by a constant pressure gradient, $\pgrad$, through a two-dimensional channel.
A schematic is shown in \Fref{fig:poise-schematic}.
No-slip boundary conditions are enforced at the top and bottom boundaries and periodic boundary conditions are enforced at the left and right boundaries.
The total height of the channel will be denoted by $H$.
The center of the channel is $y = 0$ and $y \in [-h, h]$ where $h = \frac{H}{2}$.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-schematic}
    \caption{Schematic of Poiseuille flow; no-slip boundary conditions are enforced at the top and bottom boundaries, and periodic boundary conditions are enforced at the left and right boundaries.}
    \label{fig:poise-schematic}
\end{figure}

Unless otherwise stated, all of the Poiseuille flow simulations were computed on a $32 \times 64$ lattice for 25000 timesteps (although the length, $L$, seems small, the effect is negligible as the east-west boundary conditions are periodic).
In reference to computational time, each of the simulations in this section were run on a single core of an Intel I7-860 Quad-Core 2.80GHz processor.
For the MRT relaxation matrix the free parameters were set to $s_1 = s_2 = s_4 = s_6 = 1.1$.

@@

The Bingham plastic simulations were carried out with a pressure gradient of $\pgrad = 1.0 \times 10^{-5}$, and a plastic viscosity of $\mu_p = 0.2$.
The yield stress was varied between the four different values $\tau_y = [4.0, 8.0, 12.0, 16.0] \times 10^{-5}$, and four different LBM schemes were used: (1) BGK with $m = 10^5$, (2) BGK with $m = 10^8$, (3) BGK with $m = 10^8$ and the median filter, and (4) MRT with $m = 10^8$. 
(Recall $m$ is the stress growth exponent for the Papanastasiou approximation.)
The relative $L_2$ and relative $L_{\infty}$ errors with respect to the analytical solution were computed for each simulation.
The analytical solution for Bingham plastic flow is given by:
\begin{equation} \label{eq:analytical-bingham}
u_x(y) = \begin{cases}
\frac{1}{2 \mu_p} \left(-\pgrad\right) \left[h^2 - y_{\tau}^2 \right] - \frac{\tau_y}{\mu_p} \left(h - y_{\tau}\right), & 0 \leq |y| \leq y_{\tau}, \\
\frac{1}{2 \mu_p} \left(-\pgrad\right) \left[h^2 - y^2 \right] - \frac{\tau_y}{\mu_p} \left(h - |y|\right), & y_{\tau} < |y| \leq h,
\end{cases}
\end{equation}
\noindent where $y_{\tau} = -\tau_y / \pgrad$ is the point at which the fluid yields.

The results are presented in \Fref{tab:poise-bing}.
The Reynold's number was computed by $Re = \frac{\rho U H}{\mu_p}$, where $U$ is the maximum velocity given by the analytical solution and $H = 2h$ is the total height of the channel.
The Bingham number was computed by $Bn = \frac{\tau_y H}{\mu_p U}$.

@@

As has been reported previously, for the BGK collision operator--regardless of whether or not median entropic filtering is used--using a stress growth exponent of $m = 10^5$ is more accurate with respect to the analytical solution than using a stress growth exponent of $m = 10^8$~\cite{chen2014simulations}.
A larger stress growth exponent results in a more accurate Papastasiou approximation of the true Bingham plastic constitutive model, however, it leads to more noise that degrades the numerical solution for the BGK collision operator.
Upon inspection of \Fref{tab:poise-bing}, it does not appear that entropic median filtering helps dissipate the noise that occurs as a result of using $m = 10^8$ for Bingham plastic Poiseuille flow.
In fact, the median filter resulted in a less accurate solution in nearly all cases and rendered the solution altogether useless for the higher yield stress flows (relative errors of 25-75\%).

The BGK collision operator using a stress growth exponent of $m = 10^5$ was in general the fastest.
This can be attributed to the fact that a smoother approximation of the Bingham plastic constitutive model would lead to a solution for $\mu_{app}$ converging with less iterations.
The BGK collision operator using a stress growth exponent of $m = 10^5$ also performed acceptably in terms of accuracy for lower yield stress flows, however, for $\tau_y = 12 \times 10^{-5}$ and $16 \times 10^{-5}$ the relative $L_\infty$ errors were 8.9\% and 19\%, respectively--larger than what would be considered acceptable for most engineering applications.

The LBM approximations are plotted with the analytical solution for the BGK collision operator using $m = 10^5$ and $m = 10^8$ in \Fref{fig:bing-analyt-bgk-5} and \Fref{fig:bing-analyt-bgk-8}, respectively.
\Fref{fig:bing-analyt-bgk-5} shows that the error for the BGK model with $m = 10^5$ is not due to noise, but instead the inaccuracy of the Papastasiou approximation with a lower stress growth exponent.
\Fref{fig:bing-analyt-bgk-8} shows that the error for the BGK model with $m = 10^8$ is due to numerical noise.
In order to better understand why median filtering did not eliminate the noise in high yield stress flows, but instead exacerbate the problem, it is necessary to investigate what is happening at the particle distribution scale.

@@

\Fref{fig:feq-vs-f_bgk} compares particle distributions to their respective quasiequilbiriums.
It can be seen that in the BGK, $m = 10^8$ solution, the noise due to nonlinear constitutive relation pollutes the quasiequilibrium values.
Because the noise makes its way into the quasiequilibrium values, entropic median filtering does not help dampen the noise but instead contracts the particle distributions closer to the quasiequilibrium noise--explaining why the median filtered results were less accurate for Bingham plastic Poiseuille flow.
In order to ensure that this phenomena was a side effect of all entropic filtering and not just entropic median filtering with $n_s = 2.7$, entropic filtering was tested with median filtering and Ehrenfests' regularization for $n_s = 2.7$ and for $\Delta S$ thresholds of $\delta = 10^{-4}$ and $\delta = 10^{-5}$.
All filtering schemes suffered from approximately the same level of noise.

%MG: reformat these images so that text is more readable
\begin{figure}
\centering
\begin{tabulary}{\linewidth}{c c}
\includegraphics[width=2.1in]{figs/poise-bingham/bgk-5/feq-vs-f_5.png}
&
\includegraphics[width=2.1in]{figs/poise-bingham/bgk-5/feq-vs-f_8.png}
\\
\includegraphics[width=2.1in]{figs/poise-bingham/bgk-8/feq-vs-f_5.png}
&
\includegraphics[width=2.1in]{figs/poise-bingham/bgk-8/feq-vs-f_8.png}
\end{tabulary}
\caption{Particle distributions in the 5 direction (left) and 8 direction (right) compared to their respective quasiequilibriums. The top two plots are for the BGK with $m = 10^5$, the bottom two are for the BGK with $m = 10^8$.}
\label{fig:feq-vs-f_bgk}
\end{figure}

On average, the LBM model using the MRT collision operator was 2.9 times slower than the fastest model for the given material parameters.
However, the increased computational time can be justified for the MRT collision operator because it did not suffer from the same level of noise as the BGK collision operator when using the more accurate approximation of the Bingham plastic model, i.e. when $m = 10^8$. 
The MRT collision operator was therefore the most accurate solution for all cases.
\Fref{fig:feq-vs-f_mrt} compares particle distributions to quasiequilibrium for the MRT model; in contrast to the BGK with $m = 10^8$, there is no noise in the quasiequilibrium distributions.

\begin{figure}
	\centering
    \begin{tabulary}{\linewidth}{c c}
        \includegraphics[width=2.1in]{figs/poise-bingham/mrt/feq-vs-f_5.png}
        &
        \includegraphics[width=2.1in]{figs/poise-bingham/mrt/feq-vs-f_8.png}
    \end{tabulary}
    \caption{Particle distributions in the 5 direction (left) and 8 direction (right) compared to their respective quasiequilibriums for the MRT with $m = 10^8$.}
    \label{fig:feq-vs-f_mrt}
\end{figure}

A natural question would be to ask ``if the strength of the MRT collision operator is damping out the ghost modes, then is the noise for the BGK with $m = 10^8$ (with and without median filtering) indeed due to the ghost modes?''.
\Fref{fig:epsilon} shows a measure of the nonequilibrium $\epsilon$ moment with respect to time for each of the collision schemes and was calculated using:
\begin{equation}
\frac{||\boldsymbol{\epsilon} - \boldsymbol{\epsilon}^{eq}||_2}{||\boldsymbol{\epsilon}^{eq}||_2}
\end{equation}
\noindent where $\epsilon_j = 4f_0(\pos_j) - 2f_1(\pos_j) - 2f_2(\pos_j) - 2f_3(\pos_j) - 2f_4(\pos_j) + f_5(\pos_j) + f_6(\pos_j) + f_7(\pos_j) + f_8(\pos_j)$ is related to the square of the energy, and $\pos_j = (16, j)$ for $j = 1, 2, ..., 64$, i.e. $\pos_j$ is taken at 16 nodes in from the left in the x-direction and for the full height of the channel in the y-direction.
The values measured for the BGK with $m = 10^8$, with and without median filtering, were significantly higher than for either the MRT or BGK with $m = 10^5$ collision schemes--which suggests that the numerical noise the BGK with $m = 10^8$ suffered from was in part due to the $\epsilon$ moment.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-bingham/epsilon}
    \caption{Evolution of relative $L_2$, $\epsilon^{neq}$ with time. The norm of the $\epsilon^{neq}$ across the height of the channel is an indicator of an increase in oscillations at the lattice level due to the $\epsilon$ moment.}
    \label{fig:epsilon}
\end{figure}

\Fref{fig:qx} shows a measure of the nonequilibrium $q_x$ moment, or energy flux in the x-direction, with respect to time for each of the collision schemes and was calculated using:
\begin{equation}
    \frac{||\mathbf{q_x} - \mathbf{q_x}^{eq}||_2}{||\mathbf{q_x}^{eq}||_2}
\end{equation}
\noindent where $q_{x_j} = - 2f_1(\pos_j) + 2f_3(\pos_j) + f_5(\pos_j) - f_6(\pos_j) - f_7(\pos_j) + f_8(\pos_j)$ and $\pos_j$ were taken across the height of the channel in the same manner as with the $\epsilon$ moment.
The values measured for the BGK with $m = 10^8$ and median filtered were consistently greater than any other collision scheme.
The $q_x$ moment was likely the cause of the significant error and noise in the BGK with $m = 10^8$ and median filtering solution.
As expected, the $q_x^{neq}$ relative $L_2$ for the BGK with $m = 10^8$ was slightly greater than the BGK with $m = 10^5$, and the $q_x^{neq}$ relative $L_2$ was negligible for the MRT collision scheme.
These results are consistent with the solution noise and error.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-bingham/qx}
    \caption{Evolution of relative $L_2$, $q_x^{neq}$ with time. The norm of the $q_x^{neq}$ across the height of the channel is an indicator of an increase in oscillations at the lattice level due to the $q_x$ moment.}
    \label{fig:qx}
\end{figure}

It is likely that the nonequilibrium $\epsilon$ and $q_x$ moment cause oscillations at the lattice level that have a cumulative effect in time.
\Fref{fig:epsilon-cumulative} and \Fref{fig:qx-cumulative} show the cumulative measures of the nonequilibrium moments in time.
\Fref{fig:qx-cumulative} suggests that the $q_x^{neq}$ dominates and is the primary source of oscillations.
If an LBM collision scheme is to be developed for simulating high yield stress flows that is more accurate than the BGK with $m = 10^5$ and more computationally efficient than the MRT, it should focus on a means of dampening the nonequilibrium energy flux moments, namely $q_x$ and $q_y$.

\begin{figure}
    \includegraphics[width=\linewidth]{figs/poise-bingham/epsilon_cumulative}
    \caption{Cumulative relative $L_2$, $\epsilon^{neq}$ with time. Oscillations can have a positive effect on each other, building up. The cumulative relative $L_2$, $\epsilon^{neq}$ is a measure of how much oscillations due to the $\epsilon$ moment may have been building in time.}
    \label{fig:epsilon-cumulative}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{figs/poise-bingham/qx_cumulative}
    \caption{Cumulative relative $L_2$, $q_x^{neq}$ with time. Oscillations can have a positive effect on each other, building up. The cumulative relative $L_2$, $q_x^{neq}$ is a measure of how much oscillations due to the $q_x$ moment may have been building in time.}
    \label{fig:qx-cumulative}
\end{figure}

>>>>

Analytical solutions rarely exist for even the simplest non-Newtonian flows because of the complexity that a nonlinear constitutive relationship entails.
It is generally more practical to approximate non-Newtonian flows using numerical methods. %MG: ask John about the following pargraph...
However, the nonlinear constitutive equation--typically of the form \begin{equation} \label{eq:non-newtonian}
\tau = \mu_{app}(\dot{\gamma}) \dot{\gamma},
\end{equation}\noindent where the apparent viscosity, $\mu_{app}$ is a function of the strain rate--results in certain challenges for numerical methods as well.
Determining the apparent viscosity and strain-rate of a flow will often require an iterative solution, a general Picard-type algorithm is:

@@

A criteria for introducing artificial dissipation that has been used successfully ~\cite{gorban2014enhancement,brownlee2006stabilization,brownlee2007stability,brownlee2008nonequilibrium} is to define a threshold, $\theta$, such that dissipation is added when:
\begin{equation} \label{eq:ds-threshold}
  \Delta S(\pos, t) > \theta.
\end{equation}
A potential drawback of defining a threshold \textit{a priori} is that in order to ensure the model still retains some physical integrity, only a small number of sites can have artificial dissipation added.
If the threshold is too low, too many sites may have dissipation added.
If the threshold is too high, a stable solution may not be achieved.
The threshold can be determined on a case-by-case basis through trial-and-error or by a preliminary analysis.
In the current work, the criteria that is used for determining whether dissipation should be added is a combination of \eqref{eq:ds-threshold} and the following:

\begin{equation} \label{eq:dissipation-criteria}
  \Delta S(\pos, t) > \overline{\Delta S} + n_\sigma \cdot \sigma_{\Delta S},
\end{equation}

\noindent where $\overline{\Delta S}$ and $\sigma_{\Delta S}$ are the mean and standard deviation of $\Delta S$, respectively--both are calculated using values over the domain for the current timestep--and $n_\sigma$ is the number of standard deviations greater than $\overline{\Delta S}$ that $\Delta S$ must be before dissipation is added.
The number of standard deviations, $n_\sigma$, is chosen \textit{a priori}.
The criteria described in \eqref{eq:dissipation-criteria} has the advantage that one does not need to determine \textit{a priori} what constitutes ``far'' from quasiequilibrium, but instead can think in terms of what would be the maximum percentage of sites one would want artificial dissipation to be added to.
A disadvantage of \eqref{eq:dissipation-critera} is that if $\overline{\Delta S}$ and $n_\sigma$ are both small then it is possible for dissipation to be added when $\Delta S(\pos, t)$ is low and artifical dissipation is unnecessary.
By required that both \eqref{eq:ds-theshold} and \eqref{eq:dissipation-criteria} be met before artifical dissipation is added, there is the potential for \eqref{eq:ds-threshold} and \eqref{eq:dissipation-criteria} to compensate for the each other's disadvantage.

@@

In order to solve for the apparent viscosity, $\mu_{app}$, using the Picard-type algorithm outlined in \Fref{sec:lbm-for-nnf}, one must calculate the strain-rate, $\dot{\gamma}$.
The strain-rate is given by the second invariant of the strain-rate tensor, $\srtensor$, i.e.:
\begin{equation} \label{eq:second-invariant}
\dot{\gamma} = \sqrt{2 \sum_{\alpha,\beta=1}^{2} \srtensor \srtensor}.
\end{equation}

@@

A numerical study was carried out in order to investigate the suitability of different LBM collision schemes for simulating non-Newtonian flows in terms of accuracy, numerical stability and computational efficiency.
For all simulations presented in this section, the apparent viscosity is approximated using its corresponding constitutive relationship and the Picard-type algorithm outlined in \Fref{sec:lbm-for-nnf} with the maximum number of iterations set at 15 and the convergence criteria:
\begin{equation} \label{eq:mu-app-conv}
\frac{\left|\mu_{app}^{k+1} - \mu_{app}^{k}\right|}{\mu_{app}^{k}} < 1.0 \times 10^{-6}.
\end{equation}
All BGK-BRT collision schemes use $\omega \in [0.05, 1.995]$ as the bounds on the collision frequency.
All simulations with artificial dissipation use the median filter with $\theta = 1.0 \times 10^{-6}$ and $n_{\sigma} = 2.7$ where $\theta$ is the threshold that $\Delta S(\pos, t$ must exceed before dissipation is added \eqref{eq:ds-threshold} and $n_{\sigma}$ is the number of standard deviations greater than $\overline{\Delta S}$ that $\Delta S(\pos, t)$ must be before dissipation is added \eqref{eq:dissipation-criteria}.
All numerical values given in this section are in lattice units unless otherwise stated.

\newcommand{\pgrad}{\frac{\partial p}{\partial x}}
\newcommand{\unorm}{\mathbf{\hat{n}}}

@@

Poiseuille flow is a useful benchmark because analytical solutions exist for both Bingham plastic and power-law fluids.
Poiseuille flow is driven by a constant pressure gradient, $\pgrad$, through a two-dimensional channel.
A schematic is shown in \Fref{fig:poise-schematic}.
No-slip boundary conditions are enforced at the top and bottom boundaries with a wall velocity of zero so that $\mvel \times \unorm = 0$ where $\unorm$ is the unit normal vector to the boundary. 
Periodic boundary conditions are enforced at the left and right boundaries.
The total height of the channel is denoted by $H$.
The center of the channel is $y = 0$ and $y \in [-h, h]$ where $h = \frac{H}{2}$.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-schematic}
    \caption{Schematic of Poiseuille flow; no-slip boundary conditions are enforced at the top and bottom boundaries, and periodic boundary conditions are enforced at the left and right boundaries.
The size of the domain is $L = 32$  and $H = 64$.}
    \label{fig:poise-schematic}
\end{figure}

Unless otherwise stated, all of the Poiseuille flow simulations were computed on a $32 \times 64$ lattice for 25000 timesteps.
The lattice size was chosen to be sufficiently fine for accuracy (although the length, $L = 32$, seems small, the effect is negligible as the east-west boundary conditions are periodic) and the number of time steps were set high enough to ensure a steady-state would be reached.
In reference to computational time, each of the simulations in this section were run on a single core of an Intel I7-860 Quad-Core 2.80GHz processor.
For the MRT relaxation matrix the free parameters were set to $s_1 = s_2 = s_4 = s_6 = 1.1$, which follows the recommendation of~\cite{lallemand2000theory} for reasons of stability, and has been successfully used to simulate non-Newtonian flows in the past~\cite{fallah2012multiple,grasinger2015simulation}.

@@

The Bingham plastic simulations were carried out with a pressure gradient of $\pgrad = 1.0 \times 10^{-5}$, and a plastic viscosity of $\mu_p = 0.2$.
The yield stress was varied between four different values $\tau_y = [4.0, 8.0, 12.0, 16.0] \times 10^{-5}$, and four different LBM schemes were used: (1) BGK with $m = 10^5$, (2) BGK with $m = 10^8$, (3) BGK with $m = 10^8$ and the median filter, and (4) MRT with $m = 10^8$. 
Recall that $m$ is the stress growth exponent for the Papanastasiou approximation.
The relative $L_2$ and relative $L_{\infty}$ errors with respect to the analytical solution were computed for each simulation.
The analytical solution for Poiseuille flow of a Bingham plastic fluid is given by:
\begin{equation} \label{eq:analytical-bingham}
u_x(y) = \begin{cases}
\frac{1}{2 \mu_p} \left(-\pgrad\right) \left[h^2 - y_{\tau}^2 \right] - \frac{\tau_y}{\mu_p} \left(h - y_{\tau}\right), & 0 \leq |y| \leq y_{\tau}, \\
\frac{1}{2 \mu_p} \left(-\pgrad\right) \left[h^2 - y^2 \right] - \frac{\tau_y}{\mu_p} \left(h - |y|\right), & y_{\tau} < |y| \leq h,
\end{cases}
\end{equation}
\noindent where $y_{\tau} = -\tau_y / \pgrad$ is the vertical location at which the fluid yields.

The relative $L_2$ error, relative $L_{\infty}$ error, and computation time for each simulation are presented in \Fref{tab:poise-bing}.
The Reynold's number was computed by $Re = \frac{\rho U H}{\mu_p}$, where $U$ is the maximum velocity given by the analytical solution.
The Bingham number was computed by $Bn = \frac{\tau_y H}{\mu_p U}$.

@@

As has been reported previously, for the BGK collision operator, using a stress growth exponent of $m = 10^5$ is more accurate with respect to the analytical solution than using a stress growth exponent of $m = 10^8$~\cite{chen2014simulations}.
A larger stress growth exponent results in a more accurate Papastasiou approximation of the true Bingham plastic constitutive model, however, it leads to more nonphysical oscillations that degrade the numerical solution for the BGK collision operator.
Upon inspection of \Fref{tab:poise-bing}, it does not appear that entropic median filtering helps mitigate errors that occur as a result of using $m = 10^8$ for Bingham plastic Poiseuille flow.
In fact, the median filter resulted in a less accurate solution in all cases other than the case with the lowest yield stress considered (i.e., the smoothest flow field), and rendered the solution altogether useless for the higher yield stress flows (relative errors of approximately 25-75\%).

In general, the BGK collision operator using a stress growth exponent of $m = 10^5$ had the lowest computational time.
This can be attributed to the fact that a smoother approximation of the Bingham plastic constitutive model would lead to a solution for $\mu_{app}$ converging with less iterations.
The BGK collision operator using a stress growth exponent of $m = 10^5$ experienced relatively low error for lower yield stress flows, however, for $\tau_y = 12 \times 10^{-5}$ and $16 \times 10^{-5}$ the relative $L_\infty$ errors were 8.9\% and 19\%, respectively, which is larger than what would be considered acceptable for most engineering applications.

The LBM approximations of the velocity profile across the channel (more specifically, $u_x(\pos_j, 25000)$ where $\pos_j = \begin{Bmatrix}16 & y_j\end{Bmatrix}^T$ for $j = 1, 2, ..., 64$, i.e. $\pos_j$ is taken at 16 nodes in from the left in the x-direction and for the full height of the channel in the y-direction) for the BGK collision operator using $m = 10^5$ and $m = 10^8$ are plotted with the analytical solution \eqref{eq:bingham} in \Fref{fig:bing-analyt-bgk-5} and \Fref{fig:bing-analyt-bgk-8}, respectively.
Due to the smoothness of the LBM approximation in \Fref{fig:bing-analyt-bgk-5}, one can conclude that the error for the BGK model with $m = 10^5$ is not due to nonphysical oscillations, but instead the inaccuracy of the Papastasiou approximation with a lower stress growth exponent.
In contrast, the LBM approximation in \Fref{fig:bing-analyt-bgk-8} is not smooth, which suggests that the error for the BGK model with $m = 10^8$ is due to nonphysical oscillations.

@@

In order to better understand why median filtering did not eliminate the nonphysical oscillations in high yield stress flows, but instead exacerbates the problem, it is necessary to investigate what is happening at the particle distribution scale.
\Fref{fig:feq-vs-f_bgk} compares particle distributions to quasiequilibrium.
More specifically, \Fref{fig:feq-vs-f_bgk} compares $f_i(\pos_j, 25000)$ to $f_i^{eq}(\pos_j, 25000)$ where $\pos_j = \begin{Bmatrix}16 & y_j\end{Bmatrix}^T$ for $j = 1, 2, ..., 64$, and $i = 5, 8$.
It can be seen that in the BGK, $m = 10^8$ solution, that nonphysical oscillations pollute the quasiequilibrium profile as well.
These nonphysical oscillations likely originate at the lattice scale due to the sharp gradient in the macroscopic velocity, $\mvel$, near the walls, and the sharp gradient in $\mvel$ is a result of a sharp gradient in $\mu_{app}$, namely the sharp gradient of the constitutive relationship.
Because the nonphysical oscillations make their way into the quasiequilibrium values, entropic median filtering does not help dampen the oscillations but instead contracts the particle distributions closer to the quasiequilibrium oscillations, which explains why the median filtered results were less accurate for Bingham plastic Poiseuille flow.
In order to ensure that this phenomena was a side effect of all entropic filtering, and not just entropic median filtering with $\theta = 1.0 \times 10^{-6}$ and $n_s = 2.7$, optimization was used to find the value of $\theta$ that minimized error for both median filtering and Ehrenfests' regularization.
The optimization problem was defined as follows:
\begin{align*}
  \min_\theta \hspace{0.2in} & f(\theta)\\
\text{such that} \hspace{0.2in} & \theta \in [10^{-10}, 2.0]
\end{align*}
\noindent where $f(\theta) \equiv $ the relative $L_2$ error between the LBM approximation using $m = 10^8$ and entropic filtering with a $\Delta S$ threshold of $\theta$ (used in the criteria defined in \eqref{eq:ds-threshold}).
The bounds on $\theta$, namely $\theta \in [10^{-10}, 2]$ represent $\theta$ such that 100\% of lattice sites are filtered for every time step ($\theta = 10^{-10}$) and $\theta$ such that no lattice sites are filtered for any time step ($\theta = 2.0$).
The optimization problem was solved for the cases of using median filtering and Ehrenfests' regularization using both Brent's method and the Golden Section search.
For all four cases, the optimal solution was $\theta = 2.0$ and the cost function, $f(2.0)$, was equal to the relative $L_2$ error for the BGK collision operator, $m = 10^8$ and no entropic filtering.
The results of the optimization suggest that entropic filtering, at best does not affect the accuracy of LBM, and often adversly affects the accuracy of LBM in approximating Bingham plastic Poiseuille flow, regardless of whether median filtering or Ehrenfests' regularization is used, and regardless of what the $\Delta S$ threshold, $\theta$, is set to.
Thus, entropic filtering is not effective for increasing the accuracy of Bingham plastic Poiseuille flow.

%MG: reformat these images so that text is more readable
\begin{figure}
\centering
\begin{tabulary}{\linewidth}{c c}
\includegraphics[width=2.75in]{figs/poise-bingham/bgk-5/feq-vs-f_5.png}
&
\includegraphics[width=2.75in]{figs/poise-bingham/bgk-5/feq-vs-f_8.png}
\\
\includegraphics[width=2.75in]{figs/poise-bingham/bgk-8/feq-vs-f_5.png}
&
\includegraphics[width=2.75in]{figs/poise-bingham/bgk-8/feq-vs-f_8.png}
\end{tabulary}
\caption{Particle distributions in the 5 direction (left) and 8 direction (right) compared to their respective quasiequilibriums. The top two plots are for the BGK with $m = 10^5$, the bottom two are for the BGK with $m = 10^8$.}
\label{fig:feq-vs-f_bgk}
\end{figure}

On average, the LBM model using the MRT collision operator took 2.9 times more computation time than the BGK collision operator with $m = 10^5$.
However, the increased computational time can be justified for the MRT collision operator because MRT solution did not suffer from the same level of nonphysical oscillations as the BGK collision operator when using the more accurate approximation of the Bingham plastic model, i.e. when $m = 10^8$. 
The MRT collision operator was therefore the most accurate solution for all cases.
\Fref{fig:feq-vs-f_mrt} compares particle distribution, $f_i$, profiles to their respective quasiequilbirium profiles, $f_i^{eq}$ for the MRT model; in contrast to the BGK with $m = 10^8$, there are no nonphysical oscillations in the quasiequilibrium distributions.

\begin{figure}
	\centering
    \begin{tabulary}{\linewidth}{c c}
        \includegraphics[width=2.75in]{figs/poise-bingham/mrt/feq-vs-f_5.png}
        &
        \includegraphics[width=2.75in]{figs/poise-bingham/mrt/feq-vs-f_8.png}
    \end{tabulary}
    \caption{Particle distributions in the 5 direction (left) and 8 direction (right) compared to their respective quasiequilibriums for the MRT with $m = 10^8$.}
    \label{fig:feq-vs-f_mrt}
\end{figure}

Another important question to ask is, if the strength of the MRT collision operator is damping out the ghost modes, then are the nonphysical oscillations for the BGK with $m = 10^8$ (with and without median filtering) a result of the ghost modes?.
\Fref{fig:epsilon} shows a measure of the nonequilibrium $\epsilon$ moment, $\boldsymbol{\epsilon}^{neq}$, with respect to time for each of the collision schemes, which was calculated as:
\begin{equation}
\frac{||\boldsymbol{\epsilon} - \boldsymbol{\epsilon}^{eq}||_2}{||\boldsymbol{\epsilon}^{eq}||_2}
\end{equation}
\noindent where $||\hdots||_2$ is the Euclidean norm, $\epsilon_j = 4f_0(\pos_j) - 2f_1(\pos_j) - 2f_2(\pos_j) - 2f_3(\pos_j) - 2f_4(\pos_j) + f_5(\pos_j) + f_6(\pos_j) + f_7(\pos_j) + f_8(\pos_j)$ is related to the square of the energy, and $\pos_j = \begin{Bmatrix}16 & y_j\end{Bmatrix}^T$ for $j = 1, 2, ..., 64$, i.e. $\pos_j$ is taken at 16 nodes in from the left in the x-direction and for the full height of the channel in the y-direction.
The values measured for the BGK with $m = 10^8$, with and without median filtering, were significantly higher than for either the MRT or BGK with $m = 10^5$ collision schemes, which suggests that the nonphysical oscillations the BGK with $m = 10^8$ suffered from was in part due to the $\epsilon$ moment.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-bingham/epsilon}
    \caption{Evolution of relative $L_2$ norm, $\epsilon^{neq}$ with time. The norm of the $\epsilon^{neq}$ across the height of the channel is an indicator of an increase in oscillations at the lattice level due to the $\epsilon$ moment.}
    \label{fig:epsilon}
\end{figure}

\Fref{fig:qx} shows a measure of the nonequilibrium $q_x$ moment, $\mathbf{q_x}^{neq}$, or energy flux in the x-direction, with respect to time for each of the collision schemes, which was calculated as:
\begin{equation}
    \frac{||\mathbf{q_x} - \mathbf{q_x}^{eq}||_2}{||\mathbf{q_x}^{eq}||_2}
\end{equation}
\noindent where $q_{x_j} = - 2f_1(\pos_j) + 2f_3(\pos_j) + f_5(\pos_j) - f_6(\pos_j) - f_7(\pos_j) + f_8(\pos_j)$ and $\pos_j$ were taken across the height of the channel in the same manner as with the $\epsilon$ moment.
The values measured for the BGK with $m = 10^8$ and median filtered were consistently greater than any other collision scheme.
The relative $L_2$ norm of the $\mathbf{q_x}^{neq}$ for the BGK collision operator with $m = 10^8$ and median filtering solution spent most of the simulation between approximately 50\% and 100\%, this suggests that the $q_x$ moment was likely the cause of the nonphysical oscillations and error in its solution.
As expected (if the dominant source of error in the LBM approximation for Bingham plastic Poiseuille flow is indeed due to the $q_x$ moment), the $\mathbf{q_x^{neq}}$ relative $L_2$ norm for the BGK with $m = 10^8$ was slightly greater than the BGK with $m = 10^5$, and the $\mathbf{q_x^{neq}}$ relative $L_2$ was negligible for the MRT collision scheme.
These results are consistent with solution error.

\begin{figure}
	\centering
    \includegraphics[width=\linewidth]{figs/poise-bingham/qx}
    \caption{Evolution of relative $L_2$ norm, $\mathbf{q_x}^{neq}$ with time. The norm of the $\mathbf{q_x^{neq}}$ across the height of the channel is an indicator of an increase in oscillations at the lattice level due to the $q_x$ moment.}
    \label{fig:qx}
\end{figure}
